Visual Diagnostics for Constrained Optimisation with Application to Guided Tours

### 2021-03-17 paper/zhang-cook-laa-langrene-menendez.Rmd
some changes on optim method section

Optimisation problems are ubiquitous in many areas of study. While in some cases analytical solutions can be found, the majority of problems rely on numerical methods to find the optimal solution. These numerical methods follow iterative approaches that aim at finding the optimum by progressively improving the current solution until a desirable accuracy is achieved. Although this principle seems uncomplicated, a number of challenges arise such as the possible existence of multiple maxima (local and global), constraints and noisy objective function, and the trade-off between desirable accuracy and computational burden. In addition, the optimization results might depend on the algorithm starting values, affecting the consistency of results.

@@

A large class of methods utilises the gradient information of the objective function, with the most notable one being the gradient ascent (descent) method. Although gradient optimization methods are popular, they rely on the availability of the objective function derivatives and on the complexity of the constraints. As will be shown in the next section, the independent variable, in our optimisation problem, is a matrix of projection basis and the computational time required to perform differentation on a matrix could impede the render of tour animation. Hence, gradient-based methods are not the focus of this paper and considerations will be given to derivative-free methods. Derivative-free methods, which do not rely on the knowledge of the gradient, are more generally applicable. Derivative-free methods have been developed over the years, where the emphasis is on finding, in most cases, a near optimal solution. Below will introduce two random search methods: creeping random search and simulated annealing, and a pseudo derivative search.

<!-- Later, we will present a simulated annealing optimisation algorithm, which belongs to the class of random search methods, for optimisation with the guided tour. -->

A common search scheme utilised by both derivative-free methods and gradient methods is line search. In line search methods, users are required to provide an initial estimate $x_{1}$ and, at each iteration, a search direction $S_k$ and a step size $\alpha_k$ are generated. Then one moves on to the next point following $x_{k+1} = x_k + \alpha_kS_k$ and the process is repeated until the desire convergence is reached. While gradient-based methods choose the search direction by the gradient, derivative-free methods uses local information of the objective function to determine the search direction. The choice of step size also needs considerations, as inadequate step sizes might prevent the optimisation method to converge to an optimum. An ideal step size can be chosen via finding the value of $\alpha_k \in \mathbb{R}$ that maximises $f(x_k + \alpha_kS_k)$ with respect to $\alpha_k$ at each iteration.

Several R implementations address optimization problems with both general purpose as well as task specifics solvers. The most prominent one within the general solvers is \code{optim()} in the \CRANpkg{stats} [@stats] package, which provides both gradient-based and derivative-free optimisation functions. Another general solver specialised in non-linear optimisation is \CRANpkg{nloptr} [@nloptr]. Specific solvers for simulated annealing include \code{optim(..., method  "SANN")} and package \CRANpkg{GenSA} [@gensa] that deals with more complicated objective functions. For other task specific solvers, readers are recommended to visit the relevant sections in CRAN task review on \ctv{optimisation and mathematical programming} [@crantaskreviewoptim].

>>>>

<!-- Optimisation problems are ubiquitous in many areas of study. While in some cases analytical solutions can be found, the majority of problems rely on numerical methods to find the optimal solution. These numerical methods follow iterative approaches that aim at finding the optimum by progressively improving the current solution until a desirable accuracy is achieved. Although this principle seems uncomplicated, a number of challenges arise such as the possible existence of multiple maxima (local and global), constraints and noisy objective function, and the trade-off between desirable accuracy and computational burden. In addition, the optimization results might depend on the algorithm starting values, affecting the consistency of results.  -->

@@

A large class of methods utilises the gradient information of the objective function, with the most notable one being the gradient ascent (descent) method. Although gradient optimization methods are popular, they rely on the availability of the objective function derivatives and on the complexity of the constraints. As will be shown in the next section, the independent variable, in our optimisation problem, is a matrix of projection basis and the computational time required to perform differentation on a matrix could impede the render of tour animation. Hence, gradient-based methods are not the focus of this paper and considerations will be given to derivative-free methods.

Derivative-free methods, which do not rely on the knowledge of the gradient, are more generally applicable. Derivative-free methods have been developed over the years, where the emphasis is on finding, in most cases, a near optimal solution. Below will introduce three derivative-free methods, two of which are random search methods: creeping random search and simulated annealing, and the other one is pseudo derivative search.

Random search [in practice, non-convex, non-smooth]. The original (pure) random search, Brooks, randomly evaluate points in the space regardless of the current position. Although gaurantee global convergence, as the dimensional increases, blindly sample points in the entire space would cost a long time for convergence to occur [spall, 6]. Various algorithms improve on the pure random search by either concentrating on a narrower sampling space or a different updating machanism. Creeping random search, [R.C White], is such a variation where a candidate point is generated within a neighbourhood of the current point and the current point is only updated if the candidate point makes an improvment on the function value. [no convergence guarantee]. A different random search algorithm, simulated annealing, takes a different updating procedure than creeping random search. Rather than only updating the current point when an improvement is made, simulated annealing uses a Metropolis acceptance criterion, where nonimproving candidates can still have a chance to be accepted. The convergence of simulated annealing algorithms has been widely researched [reference, reference, reference] and a global convergence is guaranteed [under what condition].

Pseudo derivative search uses a common search scheme in optimistaion: line search. In line search methods, users are required to provide an initial estimate $x_{1}$ and, at each iteration, a search direction $S_k$ and a step size $\alpha_k$ are generated. Then one moves on to the next point following $x_{k+1} = x_k + \alpha_kS_k$ and the process is repeated until the desire convergence is reached. In derivative-free methods, local information of the objective function are used to determine the search direction. The choice of step size also needs considerations, as inadequate step sizes might prevent the optimisation method to converge to an optimum. An ideal step size can be chosen via finding the value of $\alpha_k \in \mathbb{R}$ that maximises $f(x_k + \alpha_kS_k)$ with respect to $\alpha_k$ at each iteration.

<!-- Several R implementations address optimization problems with both general purpose as well as task specifics solvers. The most prominent one within the general solvers is \code{optim()} in the \CRANpkg{stats} [@stats] package, which provides both gradient-based and derivative-free optimisation functions. Another general solver specialised in non-linear optimisation is \CRANpkg{nloptr} [@nloptr]. Specific solvers for simulated annealing include \code{optim(..., method  "SANN")} and package \CRANpkg{GenSA} [@gensa] that deals with more complicated objective functions. For other task specific solvers, readers are recommended to visit the relevant sections in CRAN task review on \ctv{optimisation and mathematical programming} [@crantaskreviewoptim]. -->

### 2021-03-17 paper/zhang-cook-laa-langrene-menendez.tex
some changes on optim method section

Optimisation problems are ubiquitous in many areas of study. While in
some cases analytical solutions can be found, the majority of problems
rely on numerical methods to find the optimal solution. These numerical
methods follow iterative approaches that aim at finding the optimum by
progressively improving the current solution until a desirable accuracy
is achieved. Although this principle seems uncomplicated, a number of
challenges arise such as the possible existence of multiple maxima
(local and global), constraints and noisy objective function, and the
trade-off between desirable accuracy and computational burden. In
addition, the optimization results might depend on the algorithm
starting values, affecting the consistency of results.

@@

A large class of methods utilises the gradient information of the
objective function, with the most notable one being the gradient ascent
(descent) method. Although gradient optimization methods are popular,
they rely on the availability of the objective function derivatives and
on the complexity of the constraints. As will be shown in the next
section, the independent variable, in our optimisation problem, is a
matrix of projection basis and the computational time required to
perform differentation on a matrix could impede the render of tour
animation. Hence, gradient-based methods are not the focus of this paper
and considerations will be given to derivative-free methods.
Derivative-free methods, which do not rely on the knowledge of the
gradient, are more generally applicable. Derivative-free methods have
been developed over the years, where the emphasis is on finding, in most
cases, a near optimal solution. Below will introduce two random search
methods: creeping random search and simulated annealing, and a pseudo
derivative search.

A common search scheme utilised by both derivative-free methods and
gradient methods is line search. In line search methods, users are
required to provide an initial estimate \(x_{1}\) and, at each
iteration, a search direction \(S_k\) and a step size \(\alpha_k\) are
generated. Then one moves on to the next point following
\(x_{k+1} = x_k + \alpha_kS_k\) and the process is repeated until the
desire convergence is reached. While gradient-based methods choose the
search direction by the gradient, derivative-free methods uses local
information of the objective function to determine the search direction.
The choice of step size also needs considerations, as inadequate step
sizes might prevent the optimisation method to converge to an optimum.
An ideal step size can be chosen via finding the value of
\(\alpha_k \in \mathbb{R}\) that maximises \(f(x_k + \alpha_kS_k)\) with
respect to \(\alpha_k\) at each iteration.

Several R implementations address optimization problems with both
general purpose as well as task specifics solvers. The most prominent
one within the general solvers is \code{optim()} in the \CRANpkg{stats}
\citep{stats} package, which provides both gradient-based and
derivative-free optimisation functions. Another general solver
specialised in non-linear optimisation is \CRANpkg{nloptr}
\citep{nloptr}. Specific solvers for simulated annealing include
\code{optim(..., method  "SANN")} and package \CRANpkg{GenSA}
\citep{gensa} that deals with more complicated objective functions. For
other task specific solvers, readers are recommended to visit the
relevant sections in CRAN task review on
\ctv{optimisation and mathematical programming}
\citep{crantaskreviewoptim}.

>>>>

@@

A large class of methods utilises the gradient information of the
objective function, with the most notable one being the gradient ascent
(descent) method. Although gradient optimization methods are popular,
they rely on the availability of the objective function derivatives and
on the complexity of the constraints. As will be shown in the next
section, the independent variable, in our optimisation problem, is a
matrix of projection basis and the computational time required to
perform differentation on a matrix could impede the render of tour
animation. Hence, gradient-based methods are not the focus of this paper
and considerations will be given to derivative-free methods.

Derivative-free methods, which do not rely on the knowledge of the
gradient, are more generally applicable. Derivative-free methods have
been developed over the years, where the emphasis is on finding, in most
cases, a near optimal solution. Below will introduce three
derivative-free methods, two of which are random search methods:
creeping random search and simulated annealing, and the other one is
pseudo derivative search.

Random search {[}in practice, non-convex, non-smooth{]}. The original
(pure) random search, Brooks, randomly evaluate points in the space
regardless of the current position. Although gaurantee global
convergence, as the dimensional increases, blindly sample points in the
entire space would cost a long time for convergence to occur {[}spall,
6{]}. Various algorithms improve on the pure random search by either
concentrating on a narrower sampling space or a different updating
machanism. Creeping random search, {[}R.C White{]}, is such a variation
where a candidate point is generated within a neighbourhood of the
current point and the current point is only updated if the candidate
point makes an improvment on the function value. {[}no convergence
guarantee{]}. A different random search algorithm, simulated annealing,
takes a different updating procedure than creeping random search. Rather
than only updating the current point when an improvement is made,
simulated annealing uses a Metropolis acceptance criterion, where
nonimproving candidates can still have a chance to be accepted. The
convergence of simulated annealing algorithms has been widely researched
{[}reference, reference, reference{]} and a global convergence is
guaranteed {[}under what condition{]}.

Pseudo derivative search uses a common search scheme in optimistaion:
line search. In line search methods, users are required to provide an
initial estimate \(x_{1}\) and, at each iteration, a search direction
\(S_k\) and a step size \(\alpha_k\) are generated. Then one moves on to
the next point following \(x_{k+1} = x_k + \alpha_kS_k\) and the process
is repeated until the desire convergence is reached. In derivative-free
methods, local information of the objective function are used to
determine the search direction. The choice of step size also needs
considerations, as inadequate step sizes might prevent the optimisation
method to converge to an optimum. An ideal step size can be chosen via
finding the value of \(\alpha_k \in \mathbb{R}\) that maximises
\(f(x_k + \alpha_kS_k)\) with respect to \(\alpha_k\) at each iteration.
